The evolution of gene expression in the term placenta of viviparous mammals

### 2016-01-08 manuscript/manuscript.Rnw
discuss armadillo placenta transcriptome

The top 50 expressed genes in the term placenta of \armadillo are
shown in \cref{tab:top50_armadillo_placenta} with a list of all genes
with $\mathrm{FPKM}\ge 50$ in Supplemental
\cref{tab:all_dasypus_novemcinctus_placenta}.

>>>>

The top 50 expressed genes in the term placenta of \armadillo are
shown in \cref{tab:top50_armadillo_placenta} with a list of all genes
with $\mathrm{FPKM}\ge 50$ in Supplemental
\cref{tab:all_dasypus_novemcinctus_placenta}. The gene with the
highest expression is
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000045029",complete=TRUE)}
which is most similar to the \mouse \textit{Pga5} is expanded in
\opossum into three members which are expressed in placenta (the other
two being
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000033885",complete=TRUE)}
and
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000040431",complete=TRUE)}),
though the expansion is less than that seen \cow and \sheep
(\cref{fig:pag2_evolution_tree}). There is also high expression of
hemoglobin subunit beta,
\Sexpr{gene.expression("dasypus_novemcinctus","HBB",complete=TRUE)},
an un-annotated gene with high similarity to \human and \cow
\textit{OVOS2}
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000033320",complete=TRUE)}
(\cref{fig:ovos2_evolution_tree}), the growth factors
\Sexpr{gene.expression("dasypus_novemcinctus","IGF2",complete=TRUE)}
and
\Sexpr{gene.expression("dasypus_novemcinctus","IGFBP4",complete=TRUE)},
and the thymosin
\Sexpr{gene.expression("dasypus_novemcinctus","TMSB4X",complete=TRUE)}.
As has been seen previously, multiple components of the annexin
complex are also highly expressed, including
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000045634",complete=TRUE)}
(\textit{ANXA8L1}),
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000017399",complete=TRUE)}
(\textit{S100B}),
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000048504",complete=TRUE)}
(\textit{S100A10}), and
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000004237",complete=TRUE)}
(\textit{ANXA2}).

\begin{figure}
<<ovos2.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree["ENSDNOG00000033320",tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 subtree=c("ENSDNOG00000033320","ENSMODG00000017988"))
@ 
\caption{Expression of \textit{OVOS2} and related genes}
\label{fig:ovos2_evolution_tree}
\end{figure}
